CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-09-28 codecheck-paper.tex
edit text on principles

The workflow and variations just outlined below reflect our current
views on how a codecheck should be performed. They are not set in
stone, but we do believe the following core principles underpinour
CODECHECK:

@@

\textbf{2. Communication between humans is key.} Some code may just work
without any interaction (e.g.~Manuel's certificate); however, more often
there are hidden dependencies that need to be found, or revised software
installations. Allowing the codechecker and human to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would slow things down and
reduce the chances for community building.

\textbf{3. Credit is given to codecheckers.} (TODO Explain in a few
sentences what each means, e.g.~see 1,2) Ideally publons, or similar,
but can just be a citation to the publicly available DOI. (Community
DOIs use these.)

\textbf{4. Workflows must be auditable.} (TODO Explain in a few
sentences what each means, e.g.~see 1,2)

\textbf{5. Open by default.}  By default, unless there are strong
reasons to the contrary (e.g. clinically-sensitive data), all code and
data, both from author and codechecker, will be made freely available
when the certificate is published, following community principles.

>>>>

The workflow and variations just outlined below reflect our current
views on how a codecheck should be performed. They are not set in
stone, but we do believe the following core principles underpin our
CODECHECK:

@@

\textbf{2. Communication between humans is key.} Some code may just work
without any interaction (e.g.~Manuel's certificate TODO add cert number); 
however, more often
there are hidden dependencies that need to be found, or revised software
installations. Allowing the codechecker and human to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would slow things down and
reduce the chances for community building as well as the lessons learned
for both author and codecheckers.

\textbf{3. Credit is given to codecheckers.} The value of contributing a
CODECHECK is comparable to providing a scientific peer review. Indiviual
variations aside, it may also require a similar amount of time. Therefore,
the codechecker's activity should be publicly recorded and mentioned in the
published paper.
The public record can be realised by publishing the certificate in a 
citable form (i.e., with a DOI; this is done within the community codecheck
process), by listing codecheckers on the journal's website or, ideally, by
publishing the checks alongside peer review activities in public databases,
such as ORCID \footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}}
or Publons.

\textbf{4. Workflows must be auditable.} This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Tode could be executed at least once without critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a "one~click" step, but achieving this requires particular 
skills. It is also hard to create a sufficient level of documentation for 
third parties. That is why the possibility to communicate is more important
than perfected workflows, and why a CODECHECK is not automated.
A codechecker does \emph{not} conduct a code review.
% more content from the website, may fit better elsewhere:
% The CODECHECK is not automated on purpose: automation may (a) lead to people gaming the system, (b) hide details that eventually decrease level of certainty that a codechecker has in their assessment, and (c) reduce the understandability of instructions in the long term, which is more important than short term ease of use (see Principle 2).

\textbf{5. Open by default.}  By default, unless there are strong
reasons to the contrary (e.g., clinically-sensitive data), all code and
data, both from author and codechecker, will be made freely available, i.e., 
under open licenses, when the certificate is published. The code and data 
publication should follow community good practices.
